Effect of quality-adjusted production function inputs

### 2015-06-01 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Refactored dynamic solow residual code. Now using models more effectively.

<<Dynamic Solow Residuals>>=
# Calculate finite difference approximations of time derivatives for output and all factors of production
lag <- 1       # Desired lag for calculations. Set to n for differences across n years.
pad <- "tail"  # Desired padding for NA values

dframes <- list() # Fill with data frames for each combination of country, flavor, energy, model, and nest
for (Coun in countries){
  for (flav in flavors){

    forms <- list(noE = keep_summands(formulas[[Coun]][[flav]], -3), withE = formulas[[Coun]][[flav]])
    for (form in forms){
      denom <- summands(form, n=length(summands(form))) # time variable
      nums <- all.vars(form) [1:(length(all.vars(form))-1)] # y, k, l, and e (if present)
      df <- historicalData[[Coun]][c(denom, nums)] # columns are: iYear, y, k, l, and e (if present)
      # Calculate names for the columns of finite difference derivatives
      derivNames <- lapply(1:length(nums), function(i){paste0("d", nums[[i]], "/", "dt")}) %>% unlist
      derivCols <- lapply(1:length(nums), function(i){
        # Calculate finite difference derivative columns
        ediff(df[[i+1]], pad=pad, lag=lag) / ediff(df[[1]], pad=pad, lag=lag)
      }) %>% do.call(cbind, .) %>% as.data.frame
      names(derivCols) <- derivNames
      df <- cbind(derivCols, historicalData[[Coun]][ , c(nums, "Year")])
      e <- if (length(summands(form))==4) "withE" else "noE"
      df$flavor <- if (flav == "unadjusted") "Unadjusted" else "Quality-adjusted"
      df$energy <- if (e == "noE") "Without energy" else "With energy"
      # Add a column for 1/y * dy/dt
      df[["dydtOvery"]] <- 1 / df[[nums[[1]]]] * df[[derivNames[[1]]]]

#       model.types <- c(CD="CD")
      model.types <- c(CD="CD", CES="CES")
      for (m in model.types){
        if (m == "CD"){
          origCoeffs <- filter(AllCoeffs, model==m & energy==e & Country==Coun & flavor==flav)
          df.CD <- cbind(df, origCoeffs)
          df.CD$nest <- NA
          lambda_fd <- 
            df.CD[["dydtOvery"]] - 
            df.CD[["alpha_1"]] / df.CD[[nums[[2]]]] * df.CD[[derivNames[[2]]]] - 
            df.CD[["alpha_2"]] / df.CD[[nums[[3]]]] * df.CD[[derivNames[[3]]]]
          if (e == "withE"){
            # Subtract the energy term also
            lambda_fd <- lambda_fd - 
              df.CD[["alpha_3"]] / df.CD[[nums[[4]]]] * df.CD[[derivNames[[4]]]]
          }
          # Fraction of growth explained by Solow Residual
          solowFrac <- lambda_fd / df.CD[["dydtOvery"]]
          # Add to data frame
          df.CD$lambda_fd <- lambda_fd
          df.CD$solowFrac <- solowFrac
          # Add to the list of data frames
          dframes[[length(dframes)+1]] <- df.CD
        } else if (m == "CES"){
          if (e == "noE"){
            origCoeffs <- filter(AllCoeffs, model==m & energy==e & Country==Coun & flavor==flav)
            df.kl <- cbind(df, select(origCoeffs, -nest, -flavor, -energy))
            df.kl$nest <- "kl"
            n <- c(1,2)
            # Get the values for x1 and x2
            df.kl$x1 <- df.kl[[ nums[ unlist(n)[1]+1 ] ]] # GDP is first column, add 1 to get to the factors of production
            df.kl$x2 <- df.kl[[ nums[ unlist(n)[2]+1 ] ]]
            df.kl$dx1dt <- df.kl[[ derivNames[ unlist(n)[1]+1 ] ]]
            df.kl$dx2dt <- df.kl[[ derivNames[ unlist(n)[2]+1 ] ]]
            # Calculate derivative terms
            df.kl <- df.kl %>% 
              transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
              transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
              transform(lambda_fd = dydtOvery + 1/b * dbdtOverrho1) %>%
              transform(solowFrac = lambda_fd / dydtOvery)
            # Add to the list of data frames
            dframes[[length(dframes)+1]] <- df.kl
          } else if (e == "withE"){
            nests <- list(kle=c(1,2,3), lek=c(2,3,1), ekl=c(3,1,2))
            for (i in 1:3){
              n <- nests[i] # preserves name
              origCoeffs <- filter(AllCoeffs, model==m & energy==e & Country==Coun & flavor==flav & nest==names(n)[[1]])
              df.CESe <- cbind(df, select(origCoeffs, -nest, -flavor, -energy))
              df.CESe$nest <- names(n)[[1]]
              # Unpack names so that the formulae below are readable.
              df.CESe$x1 <- df.CESe[[ nums[ unlist(n)[1]+1 ] ]] # GDP is 1st column, add 1 to get factors of production
              df.CESe$x2 <- df.CESe[[ nums[ unlist(n)[2]+1 ] ]]
              df.CESe$x3 <- df.CESe[[ nums[ unlist(n)[3]+1 ] ]]
              df.CESe$dx1dt <- df.CESe[[ derivNames[ unlist(n)[1]+1 ] ]]
              df.CESe$dx2dt <- df.CESe[[ derivNames[ unlist(n)[2]+1 ] ]]
              df.CESe$dx3dt <- df.CESe[[ derivNames[ unlist(n)[3]+1 ] ]]
              # Calculate derivative terms
              df.CESe <- df.CESe %>% 
                transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
                transform(a = delta * b^(rho/rho_1) + (1 - delta) * x3^(-rho)) %>%
                transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
                transform(dadtOverrho = delta * b^(rho/rho_1 - 1) * dbdtOverrho1 - (1 - delta) * x3^(-(rho+1))*dx3dt) %>%
                transform(lambda_fd = dydtOvery + 1/a * dadtOverrho) %>%
                transform(solowFrac = lambda_fd / dydtOvery)
              # Add to the list of data frames
              dframes[[length(dframes)+1]] <- df.CESe
            }
          } else {
            stop (paste("Unknown e:", e))
          }
        } else {
          stop(paste("Unknown model", model))
        }
      }
    }
  }      
}
temp_data <- do.call(plyr::rbind.fill, dframes)

@@

<<Trajectory_Means_CD, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CD") %>% select(-nest), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for Cobb-Douglas models.",
               label="tab:traj.means.CD")
# xtable::xtable(filter(traj.data.means, model=="CD") %>% select(-nest),
#                caption="Average $\\Delta\\lambda$ and $\\Delta r$ for Cobb-Douglas models.",
#                label="tab:traj.means.CD", 
#                digits=6) %>% 
#   print(., 
#         caption.placement="top", 
#         size="\\tiny", 
#         include.rownames=FALSE, 
#         math.style.negative=TRUE)
@
%
These results for the Cobb-Douglas model indicate that 
we should perform the same analysis for the CES model.

@@

@@

@@

>>>>

<<Dynamic Solow Residuals>>=
# Calculate finite difference approximations of time derivatives for output and all factors of production
dynamicSR <- function(model, id, lag=1, pad="tail"){
  id_list <- strsplit(id, split = "[.]")[[1]]
  energy <- id_list[[3]]
  modelStr <- id_list[[4]]

  f <- model$formula
  denom <- summands(f, n=length(summands(f))) # time variable
  nums <- all.vars(f) [1:(length(all.vars(f))-1)] # y, k, l, and e (if present)
  df <- model$data[c(denom, nums)] # columns are: iYear, y, k, l, and e (if present)

  # Calculate names for the columns of finite difference derivatives
  derivNames <- lapply(1:length(nums), function(i){paste0("d", nums[[i]], ".d", denom)}) %>% unlist
  derivCols <- lapply(1:length(nums), function(i){
    # Calculate finite difference derivative columns
    ediff(df[[i+1]], pad=pad, lag=lag) / ediff(df[[1]], pad=pad, lag=lag)
  }) %>% do.call(cbind, .) %>% as.data.frame
  names(derivCols) <- derivNames  
  df <- cbind(derivCols, df[ , c(nums, denom)]) # puts derivatives first, then raw numbers, then time
  df[["dydtOvery"]] <- 1 / df[[ nums[[1]] ]] * df[[ derivNames[[1]] ]]
  coeffs <- naturalCoef(model)
  df <- cbind(df, coeffs)

  if (modelStr == "CD"){
    lambda_fd <- 
      df[["dydtOvery"]] - 
      df[["alpha_1"]] / df[[nums[[2]]]] * df[[derivNames[[2]]]] - 
      df[["alpha_2"]] / df[[nums[[3]]]] * df[[derivNames[[3]]]]
    if (energy == "withE"){
      # Subtract the energy term also
      lambda_fd <- lambda_fd - 
        df[["alpha_3"]] / df[[nums[[4]]]] * df[[derivNames[[4]]]]
    }
    # Fraction of growth explained by Solow Residual
    solowFrac <- lambda_fd / df[["dydtOvery"]]
    # Add to data frame
    df$lambda_fd <- lambda_fd
    df$solowFrac <- solowFrac
    return(df)
  } else if (modelStr == "CES"){
    # Assign nice variable names so that the equations below are easier to read (and debug, and maintain)
    # This code is nest-aware.
    n <- model$nest
    for (i in 1:length(model$nest)){
      df[[paste0("x", i)]] <- df[[ nums[ model$nest[i]+1 ] ]] # GDP is 1st col, add 1 to get to the factors of production
      df[[paste0("dx", i, "dt")]] <- df[[ derivNames[ model$nest[i]+1 ] ]]
    }
    if (energy == "noE"){
      # Calculate terms for the finite-difference Solow residual (lambda_fd)
      df <- df %>% 
        transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
        transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
        transform(lambda_fd = dydtOvery + 1/b * dbdtOverrho1) %>%
        transform(solowFrac = lambda_fd / dydtOvery)
      return(df)
    } else if (energy == "withE"){
      df <- df %>% 
        transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
        transform(a = delta * b^(rho/rho_1) + (1 - delta) * x3^(-rho)) %>%
        transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
        transform(dadtOverrho = delta * b^(rho/rho_1 - 1) * dbdtOverrho1 - (1 - delta) * x3^(-(rho+1))*dx3dt) %>%
        transform(lambda_fd = dydtOvery + 1/a * dadtOverrho) %>%
        transform(solowFrac = lambda_fd / dydtOvery)
      return(df)
    } else {
      stop (paste("Unknown energy:", energy))
    }
  } else {
    stop(paste("Unknown modelStr", modelStr))
  }
}

lag <- 1 # year
temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(model, id, lag=1, pad="tail"){
  df <- dynamicSR(model, id=id, lag=lag, pad=pad)
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  df$Year <- model$data[["Year"]]
  return(df)  
}) %>% do.call(plyr::rbind.fill, .)

@@

<<Trajectory_Means_CD, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CD") %>% select(-nest), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for Cobb-Douglas models.",
               label="tab:traj.means.CD")
@
%
These results for the Cobb-Douglas model indicate that 
we should perform the same analysis for the CES model.

@@

<<Trajectory_Means_CESkle, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="kle")), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($kl$)$e$ nesting.",
               label="tab:traj.means.CESkle")
@

@@

<<Trajectory_Means_CESlek, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="lek")), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($le$)$k$ nesting.",
               label="tab:traj.means.CESlek")
@

@@

<<Trajectory_Means_CESekl, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="ekl")), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($ek$)$l$ nesting.",
               label="tab:traj.means.CESekl")
@
